---
title: Predicting Breast Cancer Treatment Outcome using Machine Learning on Mutation
  Profiles
author: "Ali Heydari"
project: Bioinformatics Final Project
output:
  html_document:
    df_print: paged
section: Data Processing Portion
---


Our data is the Breast Cancer data from TCGA (*TCGA-BRCA*) available [here](https://portal.gdc.cancer.gov/projects/TCGA-BRCA) (as of Sept. 20th 2019). The data format is *Mutation Annotated File (maf)*, and we can use a <code> maftools </code> (available [here](https://bioconductor.org/packages/release/bioc/html/maftools.html)) to explore the data. First, we want to install <code> maftools </code> through R.. 

First we want to exolore the data for biological interpretations, such as top mutations etc.
For this tast, we want to use `maftools`; below are the steps to install the software.

```{r}
options(repos="https://cran.rstudio.com" )

# Install pre-req packages 
install.packages("devtools")

#Install Bioconductor dependencies.
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("ComplexHeatmap")
BiocManager::install("VariantAnnotation")
BiocManager::install("Biostrings")

```
Now we want to install maftools

```{r}
#Install maftools from github repository.
library("rlang")
library("devtools")
install_github(repo = "PoisonAlien/maftools")

```

```{r}
# load maftools
require(maftools)

laml.maf = system.file('extdata', 'TCGA.GBM.mutect.da904cd3-79d7-4ae3-b6c0-e7127998b3e6.DR-10.0.somatic.maf.gz', package = 'maftools') #path to TCGA LAML MAF file

#laml<- read.maf(maf ="TCGA.GBM.mutect.da904cd3-79d7-4ae3-b6c0-e7127998b3e6.DR-10.0.somatic.maf.gz", useAll = T, verbose = T)

getwd()
setwd("/Users/aliheydari/Box\ Sync/NRT-ICGE/Maftools\ Stuff/mutect.maf")
  
data <- read.maf(maf = "mutect.maf")

getSampleSummary(data)

plotmafSummary(maf = data, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE, titvRaw = FALSE)

#We will draw oncoplots for top ten mutated genes.
oncoplot(maf = data, top = 10, fontSize = 12)

geneCloud(input = data, minMut = 5)
```


